---
title: "Vegetation Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Login OpenEO

OpenEO is an R package used to obtain Earth Observation data. Connecting to the service is required to access this data, meaning you will need a Copernicus account. Here is the documentation link: <https://documentation.dataspace.copernicus.eu/APIs/openEO/R_Client/R.html>

```{r connection, eval=FALSE}
library(openeo)
# connect to the backend and authenticate
connection = connect(host = "https://openeofed.dataspace.copernicus.eu")
login()

# get the process collection to use the predefined processes of the back-end
p = processes()
```

## Extract vegetation index data

To retrieve vegetation data, Sentinel spectral bands can be accessed via the OpenEO package to calculate a vegetation index. A mosaic reconstruction is then performed using raw imagery, allowing you to assess the available coverage over your LOI while accounting for cloud cover.

### Cloud coverage treatment

To handle `cloud cover` using the OpenEO package, the cloud_cover option can be used to filter for images with less than a specific cloud cover percentage (set to 30% here). Additionally, the Scene Classification Layer (SCL) band (<https://brazil-data-cube.github.io/specifications/bands/SCL.html>) can be utilized to classify all pixels into several categories:

| Class                    | Pixel value |
|--------------------------|-------------|
| No Data                  | 0           |
| Satured or Defective     | 1           |
| Dark Area Pixels         | 2           |
| Cloud Shadows            | 3           |
| Vegetation               | 4           |
| Not Vegetated            | 5           |
| Water                    | 6           |
| Unclassified             | 7           |
| Cloud Medium Probability | 8           |
| Cloud High Probability   | 9           |
| Thin Cirrus              | 10          |
| Snow                     | 11          |

When utilizing the SCL band, it is recommended to retain only pixels 4, 5, 6, and 7.

### Function to extract data

The following function is used to extract the NDVI data. This function called here `mosaicVegetation` is using these parameters :

-   `sts` : stormsList object from `defStormsList` function of `StormR` package.
-   `period` : the 30 days period "before" (default) or "after" the storm.
-   `filename` : character. name of the output file.
-   `buffer_box` : numeric. buffer area around the loi (default, 0.05).
-   `cloud_cover` : numeric. % of cloud cover percentage allowed (default, 30).

Before running this function, `sts` is needed, the storm "Chido" here can be extracted like this : 
```{r message = FALSE, warning=FALSE, results='hide'}
library(StormR)

sds <- defStormsDataset(
  filename = "~/Documents/ibtracs.ALL.list.v04r01.csv",
  fields =c(names = "NAME", seasons = "SEASON", isoTime = "ISO_TIME",
            lon = "LON", lat = "LAT", msw = "USA_WIND", basin = "BASIN",
            rmw = "USA_RMW", pressure = "USA_PRES", poci = "USA_POCI"),
  basin = "SI",
  seasons = c(2023,as.numeric(format(Sys.time(),"%Y"))),
  verbose = 1
)

#get the loi
map_subunits <- sf::st_read("../data/storm/ne_10m_admin_0_map_subunits/ne_10m_admin_0_map_subunits.shp")
which(map_subunits$SUBUNIT=="Mayotte")
loiSf <- map_subunits[264,]$geometry |> sf::st_as_sf()

#get chido
st <- defStormsList(
  sds = sds,
  loi = loiSf,
  names = c("CHIDO")
)

```


Be careful, the loi can be too large for OpenEO. In that case, change manually the bbox by defining it like that in the function: 
```{r, eval = FALSE}
bbox = list(west=44.98, south=-13.01, east=45.30, north=-12.61)
```

```{r function,eval = FALSE, warning = FALSE, results = FALSE, message=FALSE}
folder_out = "../path_file" #local folder to stock

mosaicVegetation = function(sts, period = "before", filename = "ndvi_output.tif", buffer_bbox = 0.05, cloud_cover = 30) {
  
  #define bounding box
  bbox <- sf::st_bbox(getLOI(st)) + c(-buffer_bbox, -buffer_bbox, buffer_bbox, buffer_bbox)
  spatial_extent <- setNames(as.list(bbox), c("west", "south", "east", "north"))
  
  #define temporal extent
  storm_dates <- as.Date(sts@data[[1]]@obs.all$iso.time)
  
  if (period == "before") {
    start_date <- min(storm_dates) - 30 
    end_date   <- min(storm_dates)
  } else if (period == "after") {
    start_date <- max(storm_dates)
    end_date   <- max(storm_dates) + 30
  }
  
  # define data
  data = p$load_collection(
    id = "SENTINEL2_L2A",
    spatial_extent = spatial_extent,
    temporal_extent = list(start_date, end_date),
    bands = c("B04", "B08", "SCL")
    #properties = list("eo:cloud_cover" = function(x) x <= 30) #to use or no depending on the treatment
  )
  
  # create mask
  cloud_mask = p$reduce_dimension(
    data = data,
    dimension = "bands",
    reducer = function(data, context) {
      scl = data[3] 
      
      # only safe classes 
      is_safe = p$or(
        p$eq(scl, 4), # vegetated
        p$or(
          p$eq(scl, 5), # naked soil
          p$or(p$eq(scl, 6), p$eq(scl, 7))# water or unclassified
        )
      )
      
      return(p$not(is_safe))
    }
  )
  # computing ndvi
  data_only = p$filter_bands(data, bands = c("B04", "B08"))
  data_cleaned = p$mask(data = data_only, mask = cloud_mask)
  
  ndvi_cube = p$ndvi(data = data_cleaned, nir = "B08", red = "B04") #can be replace by the formula
  #manually compute ndvi
  #spectral_reduce = p$reduce_dimension(data = data, dimension = "bands",reducer = function(data,context) {
  #  b8 = data[1] #nir
  # b4 = data[2] #red
  
  #    return((b8-b4)/(b8+b4))
  # })
  
  # save data
  result = p$save_result(data = ndvi_cube, format = "GTiff")
  compute_result(graph = result, output_file = file.path(folder_out, filename))
}
```

For example, using a defined LOI, a mosaic reconstruction can be performed for two distinct periods: before and after the cyclone of interest (here, Chido 12/2024).

```{r beforeAfter, eval = FALSE}
# for the cyclone chido, need to find dates without to much clouds

# before chido
mosaicVegetation(st, period = "before", filename = "ndvi_before_chido.tif")

# after chido
mosaicVegetation(st, period = "after", filename = "ndvi_after_chido.tif")
```

### Output

Here, the NDVI mosaic reconstruction for Mayotte across both periods is presented

```{r mosaicsperso, echo=FALSE}
library(slickR)

mes_photos <- c("../images/sentinel/decemberNdvi.png","../images/sentinel/januaryNdvi.png")

slickR(obj = mes_photos, height = 400, width = "100%") + 
    settings(
    dots = TRUE, 
    autoplay = TRUE, 
    autoplaySpeed = 3000,
    slidesToShow = 2,   
    slidesToScroll = 1   
  )

```

NB : The Sentinel-2 mission's Quarterly Mosaics provide a clear, cloud-free overview of the area. Over a three-month period, the highest-quality pixels are retrieved to generate the best possible representation of the vegetation. However, in the context of a cyclone, a three-month timeframe may not accurately capture the state of the vegetation immediately before and after the event, as vegetation can recover or change rapidly in the aftermath.

The output of this function allows for the recreation of a product similar to these Quarterly Mosaics. It should be noted that full coverage of the LOI (Mayotte) may not always be available; blank (white) areas indicate that the region was entirely obscured by cloud cover throughout the selected period.
